

```{r setup package loading, warning=FALSE}
library(PostChicago)
library(RColorBrewer)
library(GenomicRanges)
```

# Setup

We analyse data from CapC11 (Mahara et al., Set2)
Only 0h and 72h are taken to start with!

```{r Directory structure}

#workdir <- 'C:/Users/a590n/Nextcloud/OE0624 Projects/Manuscripts/2025 Postchicago/Original/2_resubmission Bioinformatics_orNARGAB/analysis/application2_breadth_v2/'
workdir <- '..analysis/application1_breadth/'
dir.create(workdir)
setwd(workdir)

designDir <- "designDir"
outputDir <- "plots"
dir.create(outputDir)

chicagoOutputDir <- "results"

intDir <- 'intervals'

## Check if you have already stored the chicagoData tables:
dir(chicagoOutputDir)
```


### Read in the design files

```{r read baited_genes and rmapgr, warning=FALSE}
baited_genes <- baitmap2baited_genes(designDir, save = FALSE)
rmapgr <- rmap2rmapgr(designDir, save = FALSE)

##annotated baits:
baits_anno <- readRDS('designDir/baited_genes_expr_annotated.rds')

##anno_CapC:
capc_mahara <- read.csv('Mahara2024_SupplementaryTable2_CaptureC_results.csv',stringsAsFactors = FALSE)

##induced, stable, etc genes:
induced <- unique(capc_mahara[capc_mahara$genetype=='induced',]$genename)
stablyactive <- unique(capc_mahara[capc_mahara$genetype=='stably_active',]$genename)
repressed <- unique(capc_mahara[capc_mahara$genetype %in% c('repressed','downregulated'),]$genename)
inactive <- unique(capc_mahara[capc_mahara$genetype %in% 'stably_inactive',]$genename)
upregulated <- unique(capc_mahara[capc_mahara$genetype=='upregulated',]$genename)

```

### read in the datasets

```{r Read chicagoData}
dir(chicagoOutputDir)
cdlist <- loadCdList(
    resultsDir = chicagoOutputDir,
    baited_genes = baited_genes
)
names(cdlist)

pooledSamples <- cdlist[-grep("rep", names(cdlist))]
replicateSamples <- cdlist[grep("rep", names(cdlist))]

rm(cdlist)
```

# Basic statistics for Fig. S1

```{r plotSigIntsStats, fig.width=9, fig.height=9,warning=FALSE, fig.align='center', fig.cap=paste0('Basic statistics of Capture-C data.')}

##Keep only induced genes
induced_genes=lapply(replicateSamples[c(1,2,13,14)],FUN=function(x) x[x$baitID %in% baited_genes[baited_genes$genename %in% induced]$re_id])
plotStats(induced_genes, col = col, mode = 'log')

```


# Data integration

## Create interactions table

```{r Create Ints file, message=FALSE}
ints <- makeIntsTable(pooledSamples[c(1,6)], baited_genes,
    repscores = TRUE, LL = replicateSamples[c(1,2,13,14)],
    ngroups = 1, outfolder = outputDir,overwrite = FALSE
)
```

## Annotate with intervals

```{r read in annotation intervals, warning=FALSE}
dir(intDir)
grl <- loadGrl(intDir = intDir)

ints <- annotateInts(ints = ints, grl = grl)
write.table(ints,'plots/ints_all.txt', sep='\t',eol='\n',quote=FALSE,row.names=FALSE)
```

##Further annotations of interactions into lost and gained

```{r lineplotsWithInteractions, fig.width=8, fig.height=4, fig.cap= 'Visualization of significant interactions within lineplots. Significant interactions around Hoxb3 are shaded.'}

## select interactions detected only at 0h, these are 'lost'
lost <- ints[ints$RA_0h_score >= 5 & ints$RA_72h_score < 3 &
               ints$RA_0h_rep3_score > ints$RA_72h_rep3_score &
               ints$RA_0h_rep4_score > ints$RA_72h_rep4_score &
               ints$RA_0h_N_downsampled >= ints$RA_72h_N_downsampled*1.5 &
               ints$RA_0h_N.1_downsampled > ints$RA_72h_N.1_downsampled &
               ints$RA_0h_N.2_downsampled > ints$RA_72h_N.2_downsampled, ]$intID

## select interactions detected only at 72h, these are 'gained'
gained <- ints[ints$RA_0h_score < 3 & ints$RA_72h_score >= 5 &
               ints$RA_0h_rep3_score < ints$RA_72h_rep3_score &
               ints$RA_0h_rep4_score < ints$RA_72h_rep4_score &
               ints$RA_0h_N_downsampled*1.5 <= ints$RA_72h_N_downsampled &
               ints$RA_0h_N.1_downsampled < ints$RA_72h_N.1_downsampled &
               ints$RA_0h_N.2_downsampled < ints$RA_72h_N.2_downsampled, ]$intID

## annotate the interaction table with interaction types:
ints$clusters_refined <- "both"
ints[ints$intID %in% lost, ]$clusters_refined <- "lost"
ints[ints$intID %in% gained, ]$clusters_refined <- "gained"

# ##Just once:
write.table(ints,'plots/ints_all.txt', sep='\t',eol='\n',quote=FALSE,row.names=FALSE)
```

## Visualize significant interactions using Lineplots

# Fig. 2A

```{r}
col <- palette.colors(3)[2:3]
colintervals <- palette.colors(5)[4:5]

#pdf('plots/Fig2_example_gainedLost.pdf')

for (name in c('Hoxb3')){

  id <- baited_genes[baited_genes$genename == name]$re_id
  k <- 31
  zoom <- 500000
  ylim <- c(0, 200)
  
  par(mfrow = c(2, 2))
  plotInteractions(pooledSamples[c(1,6)], id, k, zoom,
      ylim = ylim, show.legend = FALSE,
      show.legend.intervals = FALSE,
      name = name, intervals = grl,
      colintervals = colintervals,
      rmapgr = rmapgr, col = col,
      show.legend.outside = TRUE, d = ints,
      cex.intervals = 1, colints = c('grey',rev(col))
  )

}

#dev.off()
```


# Metaprofiles


#############################################
## 1. Matrices of K27ac/K27me3
#############################################


##Making oneGeneOnePeak tables

```{r}
quantfolder <- paste0(outputDir, "/oneGeneOnePeak_allChr")

ogoplist <- makeOneGeneOnePeak(
    grl = grl, rmapgr = rmapgr, 
    ints = ints, folder = quantfolder
)

##look at the oneGeneOnePeak table:
ogoplist[[1]][1:2,]

##The tables are also saved in the quantfolder:
list.files(path = quantfolder, pattern = "oneGeneOnePeak")
```

##Making matrices

```{r examplesOneGeneOnePeak, fig.width=8, fig.height=4, fig.cap= 'Visualization of oneGeneOnePeak interactions. Interactions with H3K27me3 and H3K27ac are shaded.'}

##Load Onegeneonepeak files:

fls_ogop <- list.files(path = quantfolder, pattern = "oneGeneOnePeak")

## H3K27me3 peaks:
ogop_K27me3 <- read.delim(paste(quantfolder, 
                                fls_ogop[grep("K27me3", fls_ogop)],sep = "/"), 
                          stringsAsFactors = FALSE)
ogop_K27me3$clusters_refined <- "K27me3_interaction"

## H3K27ac peaks:
ogop_K27ac <- read.delim(
    paste(quantfolder, fls_ogop[grep("K27ac", fls_ogop)],sep = "/"), 
    stringsAsFactors = FALSE
)
ogop_K27ac$clusters_refined <- "K27ac_interaction"

############################
## Create H3K27ac matrices
############################

name <- names(grl)[grep("K27ac", names(grl))]
name

## Zoom defines the +/- distance (in restriction fragments) around the intervals
## for which the matrices will be created
zoom <- 100

## Folder in which the matrices will be stored:
resfolder <- paste0(outputDir, "/matrices_allChr/")
if (!file.exists(resfolder)) {
    dir.create(resfolder)
}

## un-normalized matrix with reads
m_reads <- getMatrix(pooledSamples,
    zoom = zoom, d = ogop_K27ac, resfolder = resfolder, type = "reads",
    norm = FALSE, name = name, rmapgr = rmapgr
)

## matrix normalized to the center of interactions
## in sample 1 (as defined per normsam) with reads
m_reads_norm <- getMatrix(pooledSamples,
    zoom = zoom, d = ogop_K27ac, resfolder = resfolder, type = "reads",
    norm = TRUE, name = name, rmapgr = rmapgr, normsam = 1
)

## matrix normalized to the center of interactions
## in sample 2 (as defined per normsam) with reads
m_reads_norm2 <- getMatrix(pooledSamples,
    zoom = zoom, d = ogop_K27ac, resfolder = resfolder, type = "reads",
    norm = TRUE, name = name, rmapgr = rmapgr, normsam = 2
)


############################
## Create H3K27me3 matrices
############################

name <- names(grl)[grep("K27me3", names(grl))]
name

## Zoom defines the +/- distance (in restriction fragments) around the intervals
## for which the matrices will be created
zoom <- 100

## Folder in which the matrices will be stored:
resfolder <- paste0(outputDir, "/matrices_allChr/")
if (!file.exists(resfolder)) {
    dir.create(resfolder)
}

## un-normalized matrix with reads
m_reads_me3 <- getMatrix(pooledSamples,
    zoom = zoom, d = ogop_K27me3, resfolder = resfolder, type = "reads",
    norm = FALSE, name = name, rmapgr = rmapgr
)

## matrix normalized to the center of interactions
## in sample 1 (as defined per normsam) with reads
m_reads_norm_me3 <- getMatrix(pooledSamples,
    zoom = zoom, d = ogop_K27me3, resfolder = resfolder, type = "reads",
    norm = TRUE, name = name, rmapgr = rmapgr, normsam = 1
)

## matrix normalized to the center of interactions
## in sample 2 (as defined per normsam) with reads
m_reads_norm2_me3 <- getMatrix(pooledSamples,
    zoom = zoom, d = ogop_K27me3, resfolder = resfolder, type = "reads",
    norm = TRUE, name = name, rmapgr = rmapgr, normsam = 2
)


```


# Fig. 2E-G

##Plot K27ac/K27me3 peaks overlapping lost ints:
##Focussing on lost because gained K27me3-overlapping are so few!

```{r}

##Convenience function:

strsplit_string <- function(x, s1=NULL, s2=NULL){
##x: string or vector of strings
##s1: a regular expression before the desired string
##s2: a regular expression after the desired string
  a=x
  if (!is.null(s1)){
    a=unlist(strsplit(a,split=s1))
    a=a[1:length(a)%%2==0]
  }
  if (!is.null(s2)){
    a=unlist(strsplit(a,split=s2))
  }
  return(a)
}

##########################################
##STEP1: Extract all gained/lost ogops
##########################################

##Load ogop tables into a list ogops:

quantfolder <- paste0(outputDir, "/oneGeneOnePeak_allChr")

fls_ogop <- list.files(path = quantfolder, pattern = "oneGeneOnePeak",full.names = TRUE)
ogops <- lapply(fls_ogop,FUN=function(x) read.delim(x,stringsAsFactors = FALSE))
names(ogops)=strsplit_string(fls_ogop, s1='plots/oneGeneOnePeak_allChr/oneGeneOnePeak_',s2='.txt')

####################################################
#assign whether the interactions are lost
####################################################

for (i in 1:length(ogops)){
  ogops[[i]]$lost=FALSE
  #ogops[[i]]$gained=FALSE
}

##simply based on intIDs (normal interactions especially):

for (i in 1:length(ogops)){
  
  a=ogops[[i]]
  
  if('TRUE' %in% (a$intID %in% lost)){
     a[a$intID %in% lost,]$lost = TRUE
  }
  
  ogops[[i]]=a
}


##Based on the overlap of the interacting epigenetic peak with a lost int:

lostints=ints[ints$intID %in% lost,]
#gainedints=ints[ints$intID %in% gained,]

bed2gr <- function(bed){
##converts a bed with 3 columns to a gr
	gr=GRanges(bed[,1],IRanges(bed[,2],bed[,3]))
	if (dim(bed)[2]>3){
		names(gr)=bed[,4]
	} 
	if ('strand' %in% names(bed)){
		strand(gr)=bed$strand
	} else if ('Strand' %in% names(bed)){
		strand(gr)=bed$Strand
	}
	return(gr)
}

cols=c('seqnames_otherEnd', 'start_otherEnd', 'end_otherEnd')
ogops.mods=ogops ##keep only ogops with histone mods


##annotate all lost interactions among ogops. 
##Strategy:
##If an interacting K27ac or K27me3 peak overlaps a lost/gained interaction, we annotate it as lost/gained

for (id in unique(lostints$baitID)){
  print(id)
  b=lostints[lostints$baitID==id,]
  boe=bed2gr(b[,cols])
  og=lapply(ogops.mods, FUN=function(x) x[x$baitID==id,])
  ogoe=lapply(og, FUN=function(y) if(nrow(y)>0 && nrow(as.matrix(findOverlaps(boe, bed2gr(y[,grep('peaks',names(y))][,2:4]))))>0){
    ##get row names with lost interactions:
    rownames(y[unique(as.matrix(findOverlaps(boe, bed2gr(y[,grep('peaks',names(y))][,2:4])))[,2]),])
  })
  
  ##add info on lost interactions to the respective ogops:
  for(i in 1:length(ogoe)){
    if(!is.null(ogoe[[i]])){
      ogops.mods[[i]][ rownames(ogops.mods[[i]]) %in% ogoe[[i]] ,]$lost = TRUE
    }
  }
}


################################################
##save files:
################################################

ogops=ogops.mods
for(i in 1:length(ogops)){
  write.table(ogops[[i]], paste0(fls_ogop[[i]],'_annotated..txt') , sep='\t',eol='\n', quote=FALSE)
}

#################################################
##STEP2: Plots based on the annotated ogop files
#################################################

quantfolder <- paste0(outputDir, "/oneGeneOnePeak_allChr")

fls_ogop <- list.files(path = quantfolder, pattern = "oneGeneOnePeak",full.names = TRUE)
fls_ogop <- fls_ogop[grep('annotated..txt',fls_ogop)]
ogops <- lapply(fls_ogop,FUN=function(x) read.delim(x,stringsAsFactors = FALSE))
names(ogops)=strsplit_string(fls_ogop, s1="plots/oneGeneOnePeak_allChr/oneGeneOnePeak_",s2='.txt_annotated..txt')

##select ids for lost interactions:
lostac <- ogops[[grep('K27ac',names(ogops))]][ogops[[grep('K27ac',names(ogops))]]$lost,]
lostac <- paste(lostac$baitID,lostac$summit_re_id,sep=';')

lostk <- ogops[[grep('K27me',names(ogops))]][ogops[[grep('K27me',names(ogops))]]$lost,]
lostk <- paste(lostk$baitID,lostk$summit_re_id,sep=';')

length(lostac)
# [1] 141

###################################
##Fig. 2E-F
###################################

colaggr <- rep(col[c(1,3,4,5,6,2)], 2)
lty <- rep(c(1, 3), each = 6)
k=5
xlim=100

par(mfcol = c(2, 3))
  
gt <- 'lost'

##select based on interaction type
ids <- lostac
ac <- lapply(m_reads_norm, FUN=function(y) y[rownames(y) %in% ids,])

plotAggregatePeaks(ac,
    plotwhat = "median", k = k, xlim = c(-xlim, xlim),
    mainprefix = paste(gt, "ints K27ac\n"), ylab = "reads norm to RA0h",
    col = colaggr, lty = lty
)

##select based on interaction type
ids <- lostk
me <- lapply(m_reads_norm_me3, FUN=function(y) y[rownames(y) %in% ids,])

plotAggregatePeaks(me,
    plotwhat = "median", k = k, xlim = c(-xlim, xlim),
    mainprefix = paste(gt, "ints K27me3\n"), ylab = "reads norm to RA0h",
    col = colaggr, lty = lty
)


########################################################
##STEP3: Normalisation to compare peak width and shape
##QC for gainedac/lostac/lostk to compare their breadth!
########################################################

length(lostac)
# [1] 141
length(lostk)
# [1] 525

###################################
##Fig. 2G
###################################

colall <- palette.colors(7)[2:7]
colaggr <- rep(colall, 2)
lty <- rep(c(1, 3), each = 6)

col <- palette.colors(7)[2:3]

k=5
xlim=100

ylim=c(0,0.8)

par(mfcol = c(2, 3))

##A) Reads:
gt <- 'lost'

##select based on interaction type
ids <- lostac
ac <- lapply(m_reads_norm, FUN=function(y) y[rownames(y) %in% ids,])

plotAggregatePeaks(ac,
    plotwhat = "median", k = k, xlim = c(-xlim, xlim),
    mainprefix = paste(gt, "ints K27ac\n"), ylab = "reads norm to RA0h",
    col = colaggr, lty = lty
)

##Background normalization:
##Substract control signal:

acnorm_loss=ac[c(1,6)]
for(i in 1:nrow(ac[[1]])){
  acnorm_loss[[1]][i,]=ac[[1]][i,]-ac[[7]][i,]
  acnorm_loss[[2]][i,]=ac[[6]][i,]-ac[[12]][i,]
}

plotAggregatePeaks(acnorm_loss,
    plotwhat = "median", k = k, xlim = c(-xlim, xlim),
    mainprefix = paste(gt, "ints K27ac\n"), ylab = "reads norm to RA0h - con",
    col = col, lty = lty, ylim=ylim
)


##3) Lost interactions K27me3:

##select based on interaction type
ids <- lostk
me <- lapply(m_reads_norm_me3, FUN=function(y) y[rownames(y) %in% ids,])

plotAggregatePeaks(me,
    plotwhat = "median", k = k, xlim = c(-xlim, xlim),
    mainprefix = paste(gt, "ints K27me3\n"), ylab = "reads norm to RA0h",
    col = colaggr, lty = lty
)

##Background normalization:
##Substract control signal:

menorm_loss=me[c(1,6)]
for(i in 1:nrow(me[[1]])){
  menorm_loss[[1]][i,]=me[[1]][i,]-me[[7]][i,]
  menorm_loss[[2]][i,]=me[[6]][i,]-me[[12]][i,]
}

plotAggregatePeaks(menorm_loss,
    plotwhat = "median", k = k, xlim = c(-xlim, xlim),
    mainprefix = paste(gt, "ints K27me3\n"), ylab = "reads norm to RA0h - con",
    col = col, lty = lty, ylim=ylim
)
    
    
##Plot deltas of the signal  
    
delta_acnorm_loss=acnorm_loss[[1]]
delta_menorm_loss=menorm_loss[[1]]

for(i in 1:nrow(acnorm_loss[[1]])){
  delta_acnorm_loss[i,]=acnorm_loss[[2]][i,]-acnorm_loss[[1]][i,]
}

for(i in 1:nrow(menorm_loss[[1]])){
  delta_menorm_loss[i,]=menorm_loss[[2]][i,]-menorm_loss[[1]][i,]
}

deltalist = list(K27ac_loss=delta_acnorm_loss,
                 K27me3_loss=delta_menorm_loss)

##This is Figure 2G:
plotAggregatePeaks(deltalist,
    plotwhat = "median", k = k, 
    mainprefix = 'Delta RA72h.vs.RA.0h \n', ylab = "con-normalized deltas RA72h-RA0h"
    ,ylim=c(-0.5, 0)
)
abline(h=0,lty=2)
    
```


####################################################
##Gained/lost interactions!
####################################################

####################################################
##Now we will make aggregate profiles of all our 
##significant interactions
##those that are lost or gained only!
####################################################

```{r}

quantfolder <- paste(outputDir,'oneGeneOnePeak',sep='/')

colnames <- c('intID',  'baitID', 'otherEndID', 'seqnames_bait', 'start_bait', 'end_bait', 'seqnames_otherEnd', 'start_otherEnd', 'end_otherEnd')
ogop_all <- ints[,colnames]
ogop_all$summit_re_id=ogop_all$otherEndID
row.names(ogop_all)=ogop_all$intID
write.table(ogop_all,paste0(outputDir,'/oneGeneOnePeak_allChr/oneGeneOnePeak_allInts.txt'),
            sep='\t', eol='\n', quote=FALSE, row.names=TRUE)

##only gained/lost ints!
ogop_all <- ogop_all[ogop_all$intID %in% c(lost,gained),]
dim(ogop_all)
dim(ints)
write.table(ogop_all,paste0(outputDir,'/oneGeneOnePeak_allChr/oneGeneOnePeak_allLostGainedInts.txt'),
            sep='\t', eol='\n', quote=FALSE, row.names=TRUE)

```


##Matrices surrounding ALL lost/gained interactions:

```{r CreateInteractionMatrices, message=FALSE, warning=FALSE}
#name <- 'allInts'
name <- 'allLostGainedInts'
name

## Zoom defines the +/- distance (in restriction fragments) around the intervals
## for which the matrices will be created
zoom <- 100
d = ogop_all

## Folder in which the matrices will be stored:
resfolder <- paste0(outputDir, "/matrices_allChr/")
if (!file.exists(resfolder)) {
    dir.create(resfolder)
}

## un-normalized matrix with reads
m_reads <- getMatrix(pooledSamples,
    zoom = zoom, d = d, resfolder = resfolder, type = "reads",
    norm = FALSE, name = name, rmapgr = rmapgr
)

## matrix normalized to the center of interactions
## in sample 1 (as defined per normsam) with reads
m_reads_norm <- getMatrix(pooledSamples,
    zoom = zoom, d = d, resfolder = resfolder, type = "reads",
    norm = TRUE, name = name, rmapgr = rmapgr, normsam = 1
)

## matrix normalized to the center of interactions
## in sample 2 (as defined per normsam) with reads
m_reads_norm2 <- getMatrix(pooledSamples,
    zoom = zoom, d = d, resfolder = resfolder, type = "reads",
    norm = TRUE, name = name, rmapgr = rmapgr, normsam = 2
)

## matrix normalized to the center of interactions
## in sample 6 (RA72h) (as defined per normsam) with reads
m_reads_norm3 <- getMatrix(pooledSamples,
    zoom = zoom, d = d, resfolder = resfolder, type = "reads",
    norm = TRUE, name = name, rmapgr = rmapgr, normsam = 6
)

```

##Plot aggregate profiles of only those interactions that are lost/gained and not associated with K27me3
##Compare them to all lost-gained interactions:

#Fig. 2B-D, 2H, and S1C-D

```{r}

###################################################################
###################################################################
###################################################################
##Function for width comparison over lost/gained ints interactions
###################################################################
###################################################################
###################################################################

compareWidths <- function(type, lostnone, gainednone){
  ##type: what is the type of interactionstaht we compre? Will be in the title!
  ##lostnone: intIDs of lost interactions to consider
  ##gainednone: intIDs of gained interactions to be considered

  col <- palette.colors(7)[2:7]
  col <- col[c(1,3,4,5,6,2)]
  colaggr=rep(col,2)
  
  par(mfcol=c(2,3))
  
  ##############################
  ##Gained interactions, no K27me3
  ##############################
  
  gt <- paste('gained', type)
  
  ids=gainednone
  
  m.gained=lapply(m_reads_norm3, FUN=function(y) y[sapply(as.character(row.names(y)), FUN=function(x) x[1]) %in% ids,])
  
  plotAggregatePeaks(m.gained,
      plotwhat = "median", k = 3, xlim = c(-50, 50),
      mainprefix = paste0(gt, " norm to 72h\n"), ylab = "reads_norm",
      col = colaggr, lty = lty
  )
  
  ##Background normalization:
  ##Substract control signal:
      
  m.gained.norm=m.gained[c(1,6)]
  for(i in 1:nrow(m.gained[[1]])){
    m.gained.norm[[1]][i,]=m.gained[[1]][i,]-m.gained[[7]][i,]
    m.gained.norm[[2]][i,]=m.gained[[6]][i,]-m.gained[[12]][i,]
  }
  
  plotAggregatePeaks(m.gained.norm,
      plotwhat = "median", k = k, xlim = c(-xlim, xlim),
      mainprefix = paste(gt, "bg norm\n"), ylab = "reads norm to RA72h - con",
      col = col, lty = lty, ylim=ylim
  )
  
  
  ##############################
  ##Lost interactions, no K27me3
  ##############################
  
  gt <- paste ('lost', type)
  
  ids=lostnone
  
  m.lost=lapply(m_reads_norm, FUN=function(y) y[sapply(as.character(row.names(y)), FUN=function(x) x[1]) %in% ids,])
  
  plotAggregatePeaks(m.lost,
      plotwhat = "median", k = 3, xlim = c(-50, 50),
      mainprefix = paste0(gt," norm to 0h\n"), ylab = "reads_norm",
      col = colaggr, lty = lty
  )
  
  ##Background normalization:
  ##Substract control signal:
      
  m.lost.norm=m.lost[c(1,6)]
  for(i in 1:nrow(m.lost[[1]])){
    m.lost.norm[[1]][i,]=m.lost[[1]][i,]-m.lost[[7]][i,]
    m.lost.norm[[2]][i,]=m.lost[[6]][i,]-m.lost[[12]][i,]
  }
  
  plotAggregatePeaks(m.lost.norm,
      plotwhat = "median", k = k, xlim = c(-xlim, xlim),
      mainprefix = paste(gt, "bg norm\n"), ylab = "reads norm to RA0h - con",
      col = col, lty = lty, ylim=ylim
  )
  
  ###################################
  ##Calculate and plot deltas
  ###################################
  
  delta_gain=m.gained.norm[[1]]
  delta_loss=m.lost.norm[[1]]
     
  for(i in 1:nrow(m.gained.norm[[1]])){
    delta_gain[i,]=m.gained.norm[[2]][i,]-m.gained.norm[[1]][i,]
  }
  
  for(i in 1:nrow(m.lost.norm[[1]])){
    delta_loss[i,]=m.lost.norm[[2]][i,]-m.lost.norm[[1]][i,]
  }
  
  deltalist = list(delta_gain,
                   delta_loss)
  names(deltalist) = paste(c('gained', 'lost'), type,sep='_')
  
  ##plot deltas
  plotAggregatePeaks(deltalist,
      plotwhat = "median", k = k, 
      mainprefix = 'Delta RA72h.vs.RA.0h', ylab = "con-normalized deltas RA72h-RA0h"
      ,ylim=c(-0.35, 0.35)
  )
  abline(h=0,lty=2)
  
  ##########################
  ##flip loss for comparison
  ##########################
  
  deltalist[[grep('lost',names(deltalist))]] = -deltalist[[grep('lost',names(deltalist))]]
  
  ##plot flipped deltas
  plotAggregatePeaks(deltalist,
      plotwhat = "median", k = 5, 
      mainprefix = 'Flipped Delta RA72h.vs.RA.0h \n', ylab = "con-normalized deltas RA72h-RA0h"
      ,ylim=c(0, 0.45)
  )

}


##Run the function

type <- 'noK27me3'
lostnone <- lost[lost%in%ints[ints$K27me3_peaks_otherEnd==FALSE ,]$intID]
gainednone <- gained[gained%in%ints[ints$K27me3_peaks_otherEnd==FALSE ,]$intID]

length(lostnone)
#[1] 1687
length(gainednone)
#[1] 923
compareWidths(type,lostnone,gainednone)


type <- 'all'
length(lost)
# [1] 3609
length(gained)
# [1] 934
compareWidths(type,lost,gained)

```


##Distance seems to be an issue - let's normalize the distances!

#Fig. S1B

```{r}

#pdf('plots/Metaprofiles_LostGained_peaks_stratifiedByDistances.pdf',width=8,height=24)

lostints <- ints[ints$clusters_refined=='lost',]
gainedints <- ints[ints$clusters_refined=='gained',]

lowers <- c(1,250,500,750,1000,2000)
uppers <- c(250,500,750,1000,2000,10000)
df=data.frame(lowers,uppers)

  k=5
  xlim=100
  
  ylim=c(0,1)
  
  par(mfrow = c(3, 2))


for (j in 1:nrow(df)){
  
  lower=df[j,1]
  upper=df[j,2]
  
  lost.sel=lostints[abs(lostints$start_bait-lostints$start_otherEnd)>=lower*1000 &
                     abs(lostints$start_bait-lostints$start_otherEnd)< upper*1000, ]$intID
  gained.sel=gainedints[abs(gainedints$start_bait-gainedints$start_otherEnd)>=lower*1000 &
                     abs(gainedints$start_bait-gainedints$start_otherEnd)< upper*1000, ]$intID

##########################################
##Plot gained and lost next to each other,
##include background normalisation:
##########################################

  

    ##1) Gained interactions:
  
    gt <- paste('gained,', lower, '-',upper,'kb')
  
    ##select based on interaction type
    ids <- gained.sel
    ac <- lapply(m_reads_norm3, FUN=function(y) y[rownames(y) %in% ids,])
  
    plotAggregatePeaks(ac,
        plotwhat = "median", k = k, xlim = c(-xlim, xlim),
        mainprefix = paste(gt, "ints \n"), ylab = "reads norm to RA72h",
        col = colaggr, lty = lty, ylim=ylim
    )
    
    ##2) Lost interactions:
  
    gt <- paste('lost,', lower, '-',upper,'kb')
  
    ##select based on interaction type
    ids <- lost.sel
    ac <- lapply(m_reads_norm, FUN=function(y) y[rownames(y) %in% ids,])
  
    plotAggregatePeaks(ac,
        plotwhat = "median", k = k, xlim = c(-xlim, xlim),
        mainprefix = paste(gt, "ints \n"), ylab = "reads norm to RA0h",
        col = colaggr, lty = lty, ylim=ylim
    )
    
}

```


##Distance distributions lost/gained (Fig. S1A)

```{r}

ints$Distance.kb=abs(apply(ints[,c('start_bait','end_bait')],1,FUN='mean')-
                       apply(ints[,c('start_otherEnd','end_otherEnd')],1,FUN='mean'))/1000
                       
ints[ints$seqnames_bait!=ints$seqnames_otherEnd,]$Distance.kb = as.numeric(NA)

par(mfrow=c(1,1))

plot(density(log10(ints[ints$seqnames_bait==ints$seqnames_otherEnd & ints$clusters_refined=='gained',]$Distance.kb)),
     col='cornflowerblue',
     lwd=2,
     main='Distance distribution different interaction types',
     xlab='Distance bait-otherEnd log10([kb])')
lines(density(log10(ints[ints$seqnames_bait==ints$seqnames_otherEnd & ints$clusters_refined=='lost',]$Distance.kb)),col='orange',lwd=2)
lines(density(log10(ints[ints$seqnames_bait==ints$seqnames_otherEnd,]$Distance.kb)),col='lightgrey',lwd=2)
#abline(v=2.87,lty=3,col='cornflowerblue')
abline(v=2.9,lty=3,col='grey')
#abline(v=2.93,lty=3,col='orange')
abline(v=2.6,lty=3,col='cornflowerblue')
abline(v=2.45,lty=3,col='grey')
abline(v=4.45,lty=3,col='orange')
text(2.6,1.1,pos=2,labels=paste0(round(10^2.6,1),'kb'),col='cornflowerblue')
text(2.45,0.75,pos=2,labels=paste0(round(10^2.45,1),'kb'),col='grey')
text(2.9,1.5,pos=4,labels=paste0(round(10^2.9,1),'kb'),col='cornflowerblue')
text(4.45,0.4,pos=2,labels=paste0(round(10^4.45,1),'kb'),col='orange')
legend('topleft', legend=c('all','gained','lost'), lwd=2, col=c('lightgrey','cornflowerblue', 'orange'),bty='n',cex=0.75)

```
